### 2019-01-12 paper/conclusions-futurework.tex
Update before overleaf

\section{Conclusions and Future work}
We use lossy compression to reduce the computational overhead of
checkpointing in an adjoint computation used in seismic
inversion, a common method in seismic imaging applications whose
memory footprint commonly exceeds the available memory size in high
performance computing systems. We also developed a performance model
that computes whether or not the combination of compression and
checkpointing will outperform pure checkpointing or pure compression
in a variety of scenarios, depending on the available memory size,
computational intensity of the application, and compression ratio and
throughput of the compression algorithm. Our current result has
several limitations that we plan to address in future work:
\begin{itemize}
\item We do not discuss the accuracy of the results after decompression. This
depends on the application, compression algorithm, and affects the achievable
compression ratios. Our performance model only requires knowledge of the
compression time and ratio, and it is up to the user of this model to determine
what accuracy is needed and thus what compression ratio is realistic for their
application. TODO this is partiall addressed now. Talk about
convergence guarantees instead. 
\item ZFP only supports serial decompression. If ZFP supported
parallel decompression, our experiments would likely show a geater region in
which the combined method is faster than pure checkpointing. Furthermore, ZFP
only supports fields with up to three dimensions, while
exploiting similarities between fields at different time steps may yield a
better compression ratio. TODO update to make more relevant.
\item Our performance model is based on uniform compression ratios and
times.  However, many applications, including seismic inversion, are likely to have
initial conditions that contain little information and are easily
compressed, and the compression ratio gradually declines as the field
becomes more complex. We based our experiments on the final wave
field, which is presumably difficult to compress.
\item In comparing pure compression with pure checkpointing, we assume
that every checkpoint is compressed and decompressed. However, if the
available memory is only slightly less than the required memory, an
implementation that compresses only a subset of the checkpoints might
outperform the expectations of our model.
\item We do not discuss multi-level checkpointing, where some
checkpoints are stored on a slower, larger device. We expect
compression to be beneficial in these scenarios due to reduced data
transfer sizes.
\item TODO scheduling
\end{itemize}