### 2019-06-06 paper/errors-convergence.tex

\section{Acceptable errors and convergence}
\label{sec:errors}

Our performance model is designed to be agnostic of the specific adjoint-based 
optimization problem being solved. This is because we envision its use in a generic
checkpointing runtime that manages the checkpointed execution of the optimization
problem that accepts an acceptable error tolerance as an input parameter for each
gradient evaluation and determines whether or not compression can pay off for that
iteration, and if yes, which of the available strategies is to be used. This last question 
has previously been addressed previously in literature but in more specific contexts 
\cite{kunkel2017toward, tao2018optimizing}.

One question that arises in evaluating derivatives on grids compressed (and decompressed)
using lossy compression is the numerical stability of the computed derivatives, since errors
in neighbouring points can accumulate in the derivative rather quickly, rendering the derivatives
unusable. This question was addressed for ZFP \cite{zfp-derivatives} and SZ\cite{tao2017z}
separately.

In the context of seismic inversion, it has been shown before that the precision required
in the gradient evaluation is very low in the beginning of the optimization and
accurate gradients are not needed until the optimization is close to a minimum 
\cite{van20143d, boehm2016wavefield}. This is perhaps quite intuitive since, being far
from a minimum in the beginning, a gradient pointing in the approximate direction of the
relevant minimum is sufficient to make progress. These initial iterations could use a more
aggressive lossy compression strategy to accelerate (through compression) the progress 
towards the minimum. Once the optimization is within the vicinity of the minimum, the
gradient is required at a higher accuracy to make any progress, and this performance
model can then dynamically decide to disable compression for those iterations where
a more accurate, albeit slower, gradient evaluation is preferred.

There is also a body of work that addresses convergence guarantees of trust-region
based optimization methods in the presence of unreliable gradients. This was primarily
done for the scenario where the gradient (and sometimes the functional itself) is known
with a probability $p$.~\cite{blanchet2016convergence,cartis2017global,chen2018stochastic}
It was shown here that the convergence rate is only affected by a factor that is 
a linear function of $p$. This analytical framework could be extended to provide bounds
on the accuracy required in a particular gradient evaluation in order to guarantee a certain
convergence rate.

Both these analyses stop at the required accuracy in the gradient evaluation. This needs
to be extended to derive acceptable error tolerances in individual grid points corresponding to a specific
error bound in the overall gradient evaluation.